A Theory of Bidirectional Type Synthesis for Simple Types

### 2023-07-06 tex/sec2-outline.tex
[ tex ] Display the overview eariler
[ tex ] Do not explain the general form of bidirectional typing judgment

The construction is straightforward:
Run the mode preprocessor on the input term~$|\Gamma| \vdash t$.
If there is no synthesising mode derivation, report that $t$~is not in synthesising mode (the third outcome).
Otherwise $|\Gamma| \vdash t^{\syn}$, and we can run the bidirectional type synthesiser.
If it finds a derivation of $\Gamma \vdash t \syn A$ for some type~$A$, return a derivation of $\Gamma \vdash t : A$ (the first outcome), which is possible because the bidirectional typing~(\cref{fig:STLC-bidirectional-typing-derivations}) is \emph{sound}%
\todo{Explain the notation `$:^\dir{d}$'}
with respect to the original typing~(\cref{fig:STLC-typing-derivations}).

\begin{lemma}[Soundness]
If\/ $\Gamma \vdash t :^\dir{d} A$, then $\Gamma \vdash t : A$.
\end{lemma}

@@

\begin{lemma}[Completeness]
If\/ $|\Gamma| \vdash t^\dir{d}$ and\/ $\Gamma \vdash t : A$, then $\Gamma \vdash t :^\dir{d} A$.
\end{lemma}

>>>>

The construction is straightforward:
Run the mode preprocessor on the input term~$|\Gamma| \vdash t$.
If there is no synthesising mode derivation, report that $t$~is not in synthesising mode (the third outcome).
Otherwise $|\Gamma| \vdash t^{\syn}$, and we can run the bidirectional type synthesiser.
If it finds a derivation of $\Gamma \vdash t \syn A$ for some type~$A$, return a derivation of $\Gamma \vdash t : A$ (the first outcome), which is possible because the bidirectional typing~(\cref{fig:STLC-bidirectional-typing-derivations}) is \emph{sound}%
with respect to the original typing~(\cref{fig:STLC-typing-derivations}).

\begin{lemma}[Soundness]
If\/ $\Gamma \vdash t \syn A$, then $\Gamma \vdash t : A$.
\end{lemma}

@@

\begin{lemma}[Completeness]
If\/ $|\Gamma| \vdash t^\dir{\syn}$ and\/ $\Gamma \vdash t : A$, then $\Gamma \vdash t \syn A$.
\end{lemma}
